---
title: "Data Visualization using Plotly"
author: "Santiago Puentes Navas"
date: "7/3/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Summary

```{r cars, echo = FALSE, warning=FALSE, message = FALSE}
library(plotly)
data(mtcars)
str(mtcars)
```

For this project, I am showcasing Plotly used to compare different groups within the **mtcars** dataset. For this I will treat cylinders, engine shape (coded in vs variable) and transmission (coded in the am variable) in order to compare the result variable (mpg):

## Handling factor variables

There are some considerations done while pre-processing the dataset:

- Cylinders are simply converted using *as.factor*
- Engine shape must be converted to factor first using the aforementioned function, then setting the levels as 0 = "V-shaped" and 1 = "straight". This is indicated by the *?mtcars* command.
- Similarly, transmission needs to be encoded as it follows: 0 = "Automatic" and 1 = "Manual".

## First Boxplot: MPG vs. Number of Cylinders

```{r cylinders, echo = FALSE, message = FALSE, warning=FALSE}
plot_cyl <- plot_ly(data = mtcars, y = ~mpg, color = ~as.factor(cyl), type = "box", boxpoints = "all", jitter = 0.3)
plot_cyl
```

## Second Boxplot: MPG vs. Engine Shape

```{r enginevs, echo = FALSE, message = FALSE, warning=FALSE}
mtcars$vs = as.factor(mtcars$vs)
levels(mtcars$vs) = c("V-shaped", "straight")
plot_vs <- plot_ly(data = mtcars, y = ~mpg, color = ~vs, type = "box", boxpoints = "all", jitter = 0.3)
plot_vs
```

## Third Boxplot: MPG vs. Transmission

```{r transmission, echo = FALSE, message = FALSE, warning=FALSE}
mtcars$am = as.factor(mtcars$am)
levels(mtcars$am) = c("Automatic", "Manual")
plot_vs <- plot_ly(data = mtcars, y = ~mpg, color = ~am, type = "box", boxpoints = "all", jitter = 0.3)
plot_vs
```
